# IE - Master in Business Analytics and Big Data
Recommendation Systems

Building a Collaborative Filtering Engine

Author: Ignacio Mouawad

========================================================

# Initialization

Import libraries and Load  critics dataset.


```{r}
library(data.table)
root_path <- "C:/Users/ignac/Desktop/ie/mbd term 2/reccomendation systems/IE-Lab1/Data/"
critics <- read.csv(file.path(root_path, "critics.csv"));

```



# 1) Pearson correlation coefficient 
Calculate the Pearson correlation coefficient between Sophia and Nuria

**Expected correlation between Sophia and Nuria**: 0.3651484

```{r fig.width=7, fig.height=6}
tran_critics <- transpose(critics)
colnames(tran_critics)<-critics[,1]
rownames(tran_critics)<- colnames(critics)
tran_critics <- tran_critics[2:nrow(tran_critics),]  #this creates a transposed table of critics with users as col names and movies as row names
cor(as.numeric(tran_critics$Nuria),as.numeric(tran_critics$Sophia), use = "complete.obs", method = "pearson")
#this is to calculate the pearson corellation, aand using complete.obs to neglect the nas.
```

# 2) Compare two movie critics
Using the calculated Pearson correlation in the previous section. Compare and show the movie ratings for Sophia and Nuria on a scatter plot. How is this compared with the same ratings of Maria and Nerea on a scatter plot? What about Chris and Jim? What it means?
```{r}
tran_critics <-as.data.frame(sapply(tran_critics, as.numeric)) #we need to change them from atomic vectors to numeric
plot(tran_critics$Sophia,tran_critics$Nuria, xlab = "Sophia", ylab = "Nuria", main = "Sophia vs Nuria") #this plots Nuria vs Maria   
abline(lm(tran_critics$Nuria~tran_critics$Sophia)) #adds the linear model trend line
plot(tran_critics$Maria,tran_critics$Nerea, xlab = "Maria", ylab = "Neria", main = "Maria vs Nerea")
abline(lm(tran_critics$Nerea~tran_critics$Maria))
plot(tran_critics$Chris,tran_critics$Jim, xlab = "Chris", ylab = "Jim", main = "Chris vs Jim")
abline(lm(tran_critics$Jim~tran_critics$Chris))
```

# 3) Top Recommendations 
Return the top-5 recommended movies for Sophia Use the weighted average of every other critic's rankings.

**Expected result**:
3.7917013044215, 'The Matrix'
3.50776533175371, 'Forrest Gump'     
3.33118834864677, 'The Sixth Sense' 
3.11491825315719, 'Shakespeare in Love'
2.9124513228665, 'Blade Runner'

```{r fig.width=7, fig.height=6}
formvect <- vector() #creates empty vector to sore values of the rui-rbar*wau for each user
sumformvect <- vector()  # creates empty vector to store sum of rui-rbar*wau for each movie
absvect <- vector()
sumabsvect <- vector()
sophiaratings <- vector()#creates empty vector to get the recommended movies ratings
sophia_avg <- mean(tran_critics$Sophia, na.rm = T)



formvect <- vector()
sumformvect <- vector()
absvect <- vector()
sumabsvect <- vector()
sophiaratings <- vector()
sophia_avg <- mean(tran_critics$Sophia, na.rm = T)




for (movie in 1:ncol(tran_critics)) {
  for(user in 1:nrow(tran_critics)){
    rum <- tran_critics[movie,user]
    ravg <- mean(tran_critics[,user], na.rm = T)
    if(is.na(rum)==T| identical(tran_critics$Sophia, tran_critics[,user])){form <- 0}
    else{
      wau <- cor(tran_critics$Sophia,tran_critics[,user], use = "complete.obs")
      
      form <- (rum - ravg)*wau
      
    }
    formvect[user] <- form
    absvect[user] <- (wau) 
    
  }
  sumformvect[movie]<- sum(formvect, na.rm = T)
  sumabsvect[movie]<- sum(abs(absvect), na.rm = T)
  sophiaratings[movie] <- sophia_avg + (sumformvect[movie]/6.928035)
}

tran_critics$Sophia
tran_critics$ratings <- sophiaratings
rownames(tran_critics)<- colnames(critics)[2:ncol(critics)]
xy <-tran_critics[is.na(tran_critics$Sophia)==T,]
xy <- xy[order(xy$ratings, decreasing = T),]
xy <- xy[,c(15,ncol(xy))]
xy[2]
```

# 4) Top similar critics
Return the top 5 critics that are most similar to Sophia. 

**Expected result**:
0.9449112, 'Rachel'
0.5976143, 'Ana' 
0.5789794, 'Oriol'
0.4925922, 'Maria'
0.4273247, 'Carles'

```{r fig.width=7, fig.height=6}
similar_users <- vector()
for(user in 1:ncol(transpose(critics))){
  if(identical(tran_critics$Sophia,tran_critics[,user])){
    next
  }else{
  xrr <- (cor(tran_critics$Sophia, tran_critics[,user], use = "complete.obs", method = "pearson" ))
  }
  similar_users[user] <- xrr
}
xj <- as.data.frame(similar_users)
xj$names <- critics$User
xj <- xj[order(similar_users, decreasing = T, na.last = T),]
rownames(xj)<-NULL

xj <- xj[1:5,1:2]
rownames(xj) <- xj$similar_users
as.matrix(xj[2])
```