
```{r}
library(data.table)
library(lsa)
plays <- read.csv("C:\\Users/ignac/Desktop/ie/mbd term 2/reccomendation systems/final/play.csv")
set.seed(143)

```

```{r}
user1 <- sample(0:5,20, replace = T)
user2 <- sample(0:5,20, replace = T)
user3 <- sample(0:5,20, replace = T)
user4 <- sample(0:5,20, replace = T)
user5 <- sample(0:5,20, replace = T)
user6 <- sample(0:5,20, replace = T)
user7 <- sample(0:5,20, replace = T)
user8 <- sample(0:5,20, replace = T)
user9 <- sample(0:5,20, replace = T)
user10 <- sample(0:5,20, replace = T)

users <- data.frame(user1,user2,user3,user4,user5,  user6,user7,user8,user9,user10)
rownames(plays) <- plays[,1]
plays[,1]<-NULL
users[users==0]<-NA
rownames(users) <- rownames(plays)
users



critic1 <- sample(0:5,20, replace = T)
critic2 <- sample(0:5,20, replace = T)
critic3 <- sample(0:5,20, replace = T)
critic4 <- sample(0:5,20, replace = T)
critic5 <- sample(0:5,20, replace = T)
critic6 <- sample(0:5,20, replace = T)
critic7 <- sample(0:5,20, replace = T)
critic8 <- sample(0:5,20, replace = T)
critic9 <- sample(0:5,20, replace = T)
critic10 <- sample(0:5,20, replace = T)


top_critics <- data.frame(critic1, critic2, critic3, critic4, critic5, critic6,critic7,critic8,critic9,critic10)
rownames(top_critics) <- rownames(plays)
top_critics[top_critics==0]<-NA
top_critics

friend1 <- sample(0:5,20, replace = T)
friend2 <- sample(0:5,20, replace = T)
friend3 <- sample(0:5,20, replace = T)
friend4 <- sample(0:5,20, replace = T)
friend5 <- sample(0:5,20, replace = T)
friend6 <- sample(0:5,20, replace = T)
friend7 <- sample(0:5,20, replace = T)
friend8 <- sample(0:5,20, replace = T)
friend9 <- sample(0:5,20, replace = T)
friend10 <- sample(0:5,20, replace = T)

friends <- data.frame(friend1, friend2, friend3, friend4, friend5, friend6,friend7,friend8,friend9,friend10)
rownames(friends) <- rownames(plays)
friends[friends==0]<-NA
friends




likes_dislikes <- function(x){x[x<=2] <- -1
                              x[x==3 | is.na(x)] <- 0
                              x[x>=4] <- 1
                              return(x)}

likes_table <- as.data.frame(cbind(likes_dislikes(user1), likes_dislikes(user2), likes_dislikes(user3), likes_dislikes(user4), likes_dislikes(user5),likes_dislikes(user6), likes_dislikes(user7), likes_dislikes(user8), likes_dislikes(user9), likes_dislikes(user10)))

rownames(likes_table) <- rownames(plays)
colnames(likes_table) <- colnames(users)
likes_table

```

### Cold Start Case

## Top 5 Plays, by rating mean based on all users and critics
```{r}
playtable <- cbind(users,friends,top_critics)
mean_ratings <- sort(rowMeans(playtable, na.rm = T), decreasing = T)
as.data.frame(mean_ratings[1:5])
```

## Top 5 Plays, by rating distribution based on critics
```{r}
playtable <- cbind(users,friends,top_critics)
topperc <- sort(apply(top_critics,1,function(x) sum(x>=4, na.rm=T)/sum(x>=0, na.rm=T)),decreasing=T)
as.data.frame(topperc[1:5]) 
```
## Rating distribution based on all users and critics
```{r}
playtable <- cbind(users,friends,top_critics)
topperc <- sort(apply(playtable,1,function(x) sum(x>=4, na.rm=T)/sum(x>=0, na.rm=T)),decreasing=T)
as.data.frame(topperc[1:6]) 
```

## Top 5 Plays, by rating distribution based on critics and all users
```{r}
playtable <- cbind(users,friends,top_critics)
topperc <- sort(apply(playtable,1,function(x) sum(x>=4, na.rm=T)/sum(x>=0, na.rm=T)),decreasing=T)
top_rec <- as.data.frame(topperc[1:5]) 
colnames(top_rec) <- "rating distribution"
top_rec
```





## People who watched this play also watched this:

```{r}
topocc <- function(a){
  
  playtable <- as.data.frame(transpose(cbind(users,friends,top_critics)))
  colnames(playtable)<-rownames(plays)
  
  
  playtable <- as.data.table(playtable)[playtable[,a]!=0]
  x<- sort((nrow(playtable)+(apply(playtable[,-(as.integer(grep("Doña Rosita la soltera",colnames(playtable))))], 2,function(a){sum(a!=0, na.rm = T)})))/nrow(playtable)-1, decreasing = T)[2:6]
  x<- data.frame(x)
  colnames(x) <- sprintf("top occurances with %s",a)
  return(x)
}
#topocc("Doña Rosita la soltera")

```


## People who liked this play also liked ...

```{r}
basedonlike <- function(a){
  playtable <- as.data.frame(transpose(cbind(users,friends,top_critics)))
  colnames(playtable)<-rownames(plays)

  playtable <- playtable[playtable[,a]==4 | playtable[,a] == 5,]
  
  playtable[,a] <- NULL
  x <- sort(sapply(playtable, function(x)({mean(x, na.rm = T)})), decreasing = T)
  x <- as.data.frame(x[1:5])
  colnames(x) <- sprintf("people who liked this %s also liked this",a)
  return(x)
}

basedonlike("El Rey Leon")
```


# collaborative filtering
## top recommendations
```{r}
playtable <- cbind(users,friends,top_critics)

 


formvect <- vector()
sumformvect <- vector()
absvect <- vector()
sumabsvect <- vector()
user1ratings <- vector()
user1_avg <- mean(playtable$user1, na.rm = T)
sumabs <- sum(abs(sapply(playtable,function(a){cor(a,playtable$user1, use = "complete.obs")})))





for (play in 1:nrow(playtable)) {
  for(user in 1:ncol(playtable)){
    rum <- playtable[play,user]
    ravg <- mean(playtable[,user], na.rm = T)
    if(is.na(rum)==T| identical(playtable$user1, playtable[,play])){form <- 0}
    else{
      wau <- cor(playtable$user1,playtable[,2], use = "complete.obs")
      
      form <- (rum - ravg)*wau
      
    }
    formvect[user] <- form
    
    
  }
  sumformvect[play]<- sum(formvect, na.rm = T)
  
  user1ratings[play] <- user1_avg + (sumformvect[play]/sumabs)
}

xy <- data.table(cbind(user1ratings, rownames(plays)))
names(xy)[names(xy) == "V2"] <- "weighted rating"
xy[order(xy$user1ratings, decreasing = T),1:2][1:5]
```


## building users profiles 
## Simple Unary
### user profile
```{r}
simple_unary_user_profile <- data.frame()
for(k in 1:ncol(likes_table)){
  j<- vector()
  for(i in 1:ncol(plays)){
    j[i]<- sum(plays[,i]*likes_table[,k], na.rm = T)
  }
  simple_unary_user_profile<-rbind(simple_unary_user_profile,j)

}

colnames(simple_unary_user_profile)<-colnames(plays)
rownames(simple_unary_user_profile) <- colnames(likes_table)
simple_unary_user_profile


unary_predict_table <- data.frame()


for(i in 1:nrow(simple_unary_user_profile)){
  for(k in 1:nrow(plays)){
    plays[is.na(plays)] <- 0
    unary_predict_table[k,i]<- cosine(as.numeric(plays[k,]),as.numeric(simple_unary_user_profile[i,]))
    #unary_predict_table[1,1]<- cosine(as.numeric(plays[1,]),as.numeric(simple_unary_user_profile[1,]))
  }
}
colnames(unary_predict_table) <- rownames(simple_unary_user_profile)
unary_predict_table

likes_dislikes <- rbind(sapply(unary_predict_table, function(x){sum(x>0)}))
likes_dislikes <- rbind(likes_dislikes,sapply(unary_predict_table, function(x){sum(x<0)}))
likes_dislikes <- rbind(likes_dislikes,sapply(unary_predict_table, function(x){sum(x==0)}))
rownames(likes_dislikes) <- c("likes", "dislikes","neutral")
likes_dislikes




```


```{r}
CB_unit_weight <- data.frame()

for(k in 1:nrow(plays)){
  for(i in 1:ncol(plays)){
    CB_unit_weight[k,i] <- plays[k,i]/sum(plays[k,]) 
  }
}

unit_weight_user_profile <- data.frame()
for(k in 1:ncol(likes_table)){
  j<- vector()
  for(i in 1:ncol(CB_unit_weight)){
    j[i]<- sum(CB_unit_weight[,i]*likes_table[,k], na.rm = T)
  }
  unit_weight_user_profile<-rbind(unit_weight_user_profile,j)
  
}

colnames(unit_weight_user_profile)<-colnames(plays)
rownames(unit_weight_user_profile) <- colnames(users)
unit_weight_user_profile




unit_weight_predict_table <- data.frame()

for(i in 1:nrow(unit_weight_user_profile)){
  for(k in 1:nrow(plays)){
    unit_weight_predict_table[k,i]<- cosine(as.numeric(CB_unit_weight[k,]),as.numeric(unit_weight_user_profile[i,]))
  
  }
}
colnames(unit_weight_predict_table) <- rownames(simple_unary_user_profile)
rownames(unit_weight_predict_table) <- rownames(plays)

unit_weight_predict_table


likes_dislikes <- rbind(sapply(unit_weight_predict_table, function(x){sum(x>0)}))
likes_dislikes <- rbind(likes_dislikes,sapply(unit_weight_predict_table, function(x){sum(x<0)}))
likes_dislikes <- rbind(likes_dislikes,sapply(unit_weight_predict_table, function(x){sum(x==0)}))
rownames(likes_dislikes) <- c("likes", "dislikes","neutral")
likes_dislikes
```

## idf
```{r}
df <- sapply(plays,sum)
as.data.frame(df)
```
```{r}
idf <- log10(nrow(plays)/df)

as.data.frame(idf)
```

```{r}


idf_user_profile <- data.frame()
for(i in 1:nrow(unit_weight_user_profile)){
  idf_user_profile <- rbind(idf_user_profile,idf*unit_weight_user_profile[i,])
}
idf_user_profile
```



```{r}
idf_predict_table <- data.frame()


for(i in 1:nrow(unit_weight_user_profile)){
  for(k in 1:nrow(plays)){
    idf_predict_table[k,i]<- cosine(as.numeric(plays[k,]),as.numeric(idf_user_profile[i,]))
  
  }
}
colnames(idf_predict_table) <- rownames(simple_unary_user_profile)
rownames(idf_predict_table) <- rownames(plays)
idf_predict_table

```


```{r}

likes_dislikes <- rbind(sapply(idf_predict_table, function(x){sum(x>0)}))
likes_dislikes <- rbind(likes_dislikes,sapply(idf_predict_table, function(x){sum(x<0)}))
likes_dislikes <- rbind(likes_dislikes,sapply(idf_predict_table, function(x){sum(x==0)}))
rownames(likes_dislikes) <- c("likes", "dislikes","neutral")
likes_dislikes
```

```{r}

# top predicted Plays for user 5
xy <- data.table(cbind(idf_predict_table$user5, rownames(plays)))
names(xy)[names(xy) == "V1"] <- "weighted rating"
names(xy)[names(xy) == "V2"] <- "play"
  xy[order(xy$`weighted rating`, decreasing = T),1:2][1:5]

```

```{r}
plays[plays$classic!=0,]["classic"]
```
## Evaluation Methods
```{r}
library(DescTools)
library(rrecsys)
```



# 1) Evaluating predictions (MAE, MSE, RMSE) using Error ratings average
Combine all users predictions and observations and calculate the MAE, the MSE and the RMSE. 
Use : MAE(real,pred,na.rm.T)
What's the system status?

```{r fig.width=7, fig.height=6}
# Combine all users in one vector
all_real <- c(user1,user2,user3,user4,user5,  user6,user7,user8,user9,user10)
all_pred <- c(idf_predict_table$user1,idf_predict_table$user2,idf_predict_table$user3,idf_predict_table$user4,idf_predict_table$user5,  idf_predict_table$user6,idf_predict_table$user7,idf_predict_table$user8,idf_predict_table$user9,idf_predict_table$user10)

sprintf("Mean Absolute Error: %s", MAE(all_real,all_pred, na.rm = T))
sprintf("Mean Square Error: %s",MSE(all_real,all_pred, na.rm = T))
sprintf("Root Mean Square Error: %s",RMSE(all_real,all_pred, na.rm = T))
```
```{r fig.width=7, fig.height=6}
# Combine all users in one vector
all_real <- c(user1)
all_pred <- c(user1ratings)

sprintf("Mean Absolute Error: %s", MAE(all_real,all_pred, na.rm = T))
sprintf("Mean Square Error: %s",MSE(all_real,all_pred, na.rm = T))
sprintf("Root Mean Square Error: %s",RMSE(all_real,all_pred, na.rm = T))
```


# 2) Evaluating Users ratings average predictions, only considering RMSE
Calculate the users ratings RMSE average. Consider each user's RSME separately and then do the average of the 3 RMSEs. 
What this discrepancy means with the previous RMSE? Compare Sophia RMSE with Ivan RMSE, what it means?
```{r fig.width=7, fig.height=6}
rmse_sophia <- RMSE(sophia_real, sophia_pred, na.rm = T) ;
rmse_sophia ;
rmse_maria <- RMSE(maria_real, maria_pred, na.rm = T) ;
rmse_maria ;
rmse_ivan <- RMSE(ivan_real, ivan_pred, na.rm = T) ;
rmse_ivan;

mean(c(rmse_ivan,rmse_maria,rmse_sophia))
```